Deciding to Disclose: Pregnancy and Alcohol Misuse

### 2014-07-16 Background.tex
Working on introduction

%!TEX root = disclosure_game.tex
\section{Background}

@@

This section presents a brief overview of literature focusing on the impact
of drinking behaviour in pregnancy, and factors affecting disclosure
behaviour in the midwifery context. This is followed by a review of
literature supporting the theoretical underpinning of the modelling
approach, with particular reference to statistical decision theory.

\subsection{Alcohol, and Disclosure in the Maternity Setting}
\label{sub:alcohol_disclosure}

\subsubsection{Impact of Alcohol\label{sub:Impact-of-Alcohol}}

Distinct from stigma attached to alcohol consumption in pregnancy,
is the question of the real impact on woman and baby both in the \gls{antenatal}
period, and beyond. While the canonical example of alcohol linked
disorders is \ac{FAS}, and others on that spectrum, heavy drinking
during pregnancy has been mooted as a factor in a variety of negative
health outcomes.

The impact of moderate alcohol consumption in pregnancy is more contested.
For example, \citet{Andersen2012} examined moderate drinking in a
large Danish cohort study, finding a significant increase in the risk
of spontaneous abortion at low levels of consumption early in pregnancy.
\citet{Savitz2012} questioned the extent to which this can be interpreted
as a causal connection, noting that there is a known relationship
between absence of morning sickness, and spontaneous abortion, and
suggesting that this may explain much of the difference in risk. \citet{Kesmodel2002}
examined the relationship between alcohol consumption and still-birth,
finding that increased consumption lead to an increase in risk to
the baby, but in contrast to \citet{Andersen2012} this was significant
at \gls{term}.

Considering longer term negative outcomes, a metastudy by \citet{Latino-Martel2010}
examined the potential for maternal alcohol consumption in pregnancy
to feature as a risk factor for onset of childhood leukaemia, finding
that any alcohol consumption was associated with an increased risk
of childhood \ac{AML}, but note the rarity of the condition as a
limitation.

\citet{Huizink2006} reviewed literature looking at the impact of
moderate consumption on neurodevelopmental and cognitive outcomes,
concluding that maternal consumption can be a contributing factor
to \ac{ADHD}, and impairments to learning and memory. They subsequently
suggest that the underlying mechanism is not specific to alcohol consumption,
but a more general phenomena arising from perturbations to foetal
conditions \citep{Huizink2009}, but caution that methodological issues
in many of the studies reviewed may undermine this hypothesis.

Contrary to this, a meta-study by \citet{Gray2006} found there was
insufficient evidence to suggest any harm arising from moderate (under
1.5 UK units per day) alcohol consumption. This ties to the current
guidance from the \ac{NICE} \citep{NICE2010a}, advising that women
should avoid drinking at all in at least the first three months of
pregnancy, and no more than 1-2 units once or twice a week if they
do. In giving this advice, \ac{NICE} acknowledge that the risks to
the foetus from alcohol are a somewhat contentious subject, concluding
that the evidence of harm is inconclusive, but that this is not sufficient
to rule out the risk of negative outcomes. This tension is reflected
by earlier guidance from the \ac{RCOG} \citep{Gynecologists1996}
suggesting no evidence of harm below 15 units per week, and subsequent
criticism by \citet{Guerri1999}, who suggest that this might be interpreted
as legitimising \gls{binge}, while noting several studies indicating
adverse affects linked to even a single drink per day (e.g. \citet{Day1990}).
A subsequent \ac{RCOG} statement \citep{RoyalCollegeofObstetriciansandGynaecologists2006}
revised the recommendations to incorporate newer findings, advising
that there is no known safe threshold for drinking in pregnancy, and
highlighting binge drinking as of particular concern.

Overall, there is a distinct lack of consensus on what, and how extensive,
the effects of drinking on the immediate and long term health outcomes
are for the child.

\subsubsection{Disclosure}

The issue of disclosure is central to the model presented here, in
particular self-report by women of information that might disadvantage
them, or be expected to do so in the immediate term. In general, the
consensus is that alcohol self-reports have acceptable validity in
the research context \citep{DelBoca2000}, but do not correspond perfectly
to alternative methods. \citet{DelBoca2003} claim that the validity
is generally accepted, and suggest that the current focus lies on
what factors and processes underlie the discrepancies rather than
questioning determining their existence. In this instance, the conjecture
is that the information is in some way stigmatising; that, following
\citet{Goffman1999}, disclosure equates to revelation of the mark.
This is not immediately contentious, for example \citet{Gomberg1988}
identified stigma surrounding alcohol abusing women in particular,
an issue also highlighted by \ac{IAPT} guidance \citep{IAPT2012},
as well as a number of other studies relating response effects to
perceived negative consequences \citep{Langenbucher2001,DelBoca2000,Blair1977}.
In the maternity context, \citet{Radcliffe2011} identifies stigma
pertaining to substance misusing women amongst staff, and suggests
that this may represent a barrier to appropriate treatment; similarly,
\ac{NICE} guidance on pregnancy and complex social factors \citep{NICE2010}
recognises concern about the attitude of staff as a source of anxiety
in pregnant women who misuse substances.

Stigma, or fear of a judgemental response on the part of the practitioner
should not however be taken uncritically to explain inaccurate reporting
by patients. While recent \ac{NICE} public health guidance advocates
routine alcohol misuse screening as a part of all practice \citep{NationalInstituteforHealthandCareExcellence2010},
there is no specific policy for routine antenatal care beyond providing
information on possible impacts of alcohol consumption \citep{NICE2010a}.
\ac{NICE} guidance on pregnancy and complex social factors \citep{NICE2010}
does specifically address women who misuse alcohol, but presupposes
knowledge of the problem through medical history, or via other services.
Taken in concert with the potential for harm from even moderate alcohol
use (section \ref{sub:Impact-of-Alcohol}), this suggests that much
of the onus is on the patient to volunteer information.

Where screening is used, \citet{Kaskutas2000} note that the most
basic method, i.e. number of standard drinks consumed, can lead to
inaccurate estimates of consumption arising from inability to relate
the concept of a standard drink, to actual consumption. This is compounded
by the impact of memory effects on recall over a number of days \citep{Stockwell2004},
and a lack of consistency in the standard drink measure \citep{Turner1990}.
Alternative screening tools, for example \acs{AUDIT}, and \acs{T-ACE}
are available and have been shown to perform well in identifying problematic
levels of drinking \citep{Piccinelli1997,Bradley1998,Russell1994,Russell1996},
although the emphasis in these cases is on consumption at disordered
levels.

\citet{Prior2003} considered a different health arena (mental health
problems and GPs), with similar characteristics in terms of concealment
of medically relevant information. The central finding in this case
is that non-disclosure is not a result of stigma, but of mismatched
ontologies surrounding mental illness. Work by \citet{Alvik2005},
where the relationship between anonymity and reporting of alcohol
consumption by pregnant women was investigated, found no significant
relationship, suggesting that a fear of social judgement may not be
a dominant factor. This draws an interesting contrast with a study
by \citet{Alvik2006}, which found that contemporaneous reports of
consumption were significantly lower than those postpartum. Logistic
regression results suggest that this trend is amplified by a number
of factors, including level of alcohol consumption preceding conception,
while anxiety about foetal wellbeing during pregnancy was associated
with lower retrospective reports. Taken together with \citep{Alvik2005},
these results could be seen as conflicting, but may suggest self-stigmatisation
\citep{Watson2007}, or reflect a lack of distinction between anonymity,
and confidentiality \citep{Malvin1983}.

In summation then, there is a consensus that alcohol consumption is
generally underreported in the pregnant population, with some support
for the idea that concern about social judgement associated with stigmatisation
may be a contributing factor. Of particular interest in the wider
context of this work, is the relationship between underreporting and
consumption, i.e. that heavier drinking is associated with a greater
tendency to understate intake.

\subsubsection{Practice Implications}

Given that alcohol consumption is thought to be underreported some
consideration must be given to the implications for midwifery practice,
in terms of eliciting more accurate self-reporting. \citet{Phillips2007}
present a qualitative account of factors influencing the disclosure
of substance misuse to midwives, identifying particularly the need
to build up a rapport, potentially over a number of appointments.
This was related to continuity of care, seen as necessary by both
midwives and women for building up a trust relationship, itself a
key component of facilitating disclosure.

>>>>

%!TEX root = disclosure_game.tex
%Rewrite all of this, ditch the alcohol stuff altogether. Any really key can be woven in elsewhere (discussion)
\section{Background}

@@

This section presents a brief overview of literature focusing on the theoretical underpinning of the modelling
approach, with particular reference to statistical decision theory.

### 2014-07-16 Introduction.tex
Working on introduction

\section{Introduction}
\label{sec:intro}
%CONTRIBUTION, FUCKER!
%Also, remember to say what you're going to do for the rest of the article?

>>>>

%!TEX root = disclosure_game.tex
\section{Introduction}
\label{sec:intro}
%CONTRIBUTION, FUCKER!
%Also, remember to say what you're going to do for the rest of the article?
%CONTRIBUTION
%	Agents are awesome
%	Decision theory is super awesome
%	'nice story to hang it on'
% 	mention the importance of behaviour change

In this paper we extend previous work by \cite{GrayDissert}.

The case in favour of \ac{ABM} as a general approach has been made numerously, and elegantly REFS. As such we will not belabour the point, and instead turn to addressing some of the concerns expressed about the method. In this instance we focus on the perception of \ac{ABM} as ad hoc in nature, tending to be a reflection of the assumptions of the modeller rather than empirically or theoretically grounded \citep{Waldherr2013}. To ameliorate this concern, we draw on decision theory to produce simple rule based, learning, decision making agents and show that they are able to play a form of signalling game \citep{Kreps1987} with a basic form of intragroup information sharing. Four decision models of varying complexity, and behavioural plausibility are contrasted, by way of demonstrating the significance of the operationalisation of decision making in \ac{ABM}.

This is framed in the context of disclosure decisions, and a scenario examining drinking patterns in pregnant women which is presented in the spirit of a motivating example, rather than claimed as an accurate representation of reality. Alcohol consumption in the ante natal period is a significant issue in itself, and has been associated with many potentially negative consequences. For example, \citet{Andersen2012} report results from a large scale Danish cohort study suggesting that even low levels of consumption in early pregnancy increase the risk of spontaneous abortion, although \citet{Savitz2012} has suggested this may be attributable to a previously known link to absence of mornining sickness. Risk continues to the point of birth - \citet{Kesmodel2002} found a heightened risk to the infant - into childhood, with a metastudy by \citet{Latino-Martel2010} finding evidence on an increased risk of childhood \ac{AML} (although they suggest that the rarity of the condition is a limitation). Harm may also extend even further, and a review by \citet{Huizink2006} concluded that maternal alcohol consumption could be a contributing factor to \ac{ADHD}, and other learning impairments, but note methodological issues in a number of the papers. There is not, however, a clear consensus, with \citet{Gray2006} finding no evidence of harm below 1.5 UK units per day. In terms of official guidance, \ac{NICE} acknowledge that evidence of harm to the fetus is less than conclusive, but advise not drinking at all, or significant moderation \citep{NICE2010a}, with similar advice from the \cite{DepartmentofHealth2008}.

Turning more specifically to disclosure of alcohol use during pregnance, research is relatively sparce, although qualitative trends are reported by \citet{Phillips2007}, and \citet{Alvik2006}. The former explored factors impacting disclosure through a small case study, highlighting the need to build up rapport over several appointments; the latter compared post partum reports of consumption with contemporaneous accounts, finding apparent underreporting during pregnancy which was amplified by increased drinking. The simulation model described in this paper is able to replicate both qualitative trends, i.e. an increase in disclosure over appointments, and more honest behaviour by moderate as compared to heavier drinkers.

This scenario is of substantial independent interest, and shows the potential utility of a simulation approach in arenas where \ac{RCT} are not viable for ethical, or financial reasons. With this said, the lack of a strong quantitative evidence base against which to validate the behaviour of the model augers for caution in interpreting the results, and a necessary reminder that in this instance the focus is primarily methodological.

two fold appeal. cognitive plausibility - neuroec. inaccessible systems (litigation, concealment). fast. cheap. portable. theoretical grounding. rigor.

A game theoretic approach to generating an abstract form of the problem gives a convenient, and well known framework to reason about the processes involved in the scenario. While scenarios may well map to a plurality of games, this still allows for a principled comparison between interpretations and enforces explicit assumptions. Relating this to decision theory shifts the emphasis away from analytical equilibrium-seeking, and heightens the importance of behaviour change. Where the focus is on the behavioural processes driving a system in motion, and how they change in response to that movement, this is clearly desirable.
Fundamentally, the shift of emphasis is from the process of acquiring and inferring the information needed to make choices, to the process of decision.
Naturally, this does not preclude the incorporation of strategic refinement, since decision rules are to a great extent modular, and as demonstrated in this paper can be exchanged without altering the underpinning decision problem. In addition, rules are agnostic as to where the information used derives from, suggesting room for multi-stage processes.  As a corollary, the decision problem agents attempt to answer can change, allowing agents' behaviour in novel problems to be informed by beliefs derived under other conditions.

A key motivation for decision rules is their claim to provide an account of decision making that is behaviourally and cognitively plausible. Their mooted capability in this regard is to some extent supported by work from neuroeconomics, which aims to empirically test theories of decision making \citep{Rustichini2009}. Many key aspects common to decision rules, for example the idea that a common currency is used by the brain to compare outcomes \citep{Padoa-Schioppa2006,Padoa-Schioppa2008}, are supported by neurological findings.

Given these features, the application of decision, and game theory to \ac{ABM} is an attractive approach to computational social science, where the locus of interest is decision making. Taking a balance between the strongly biological, e.g. neural networks, and the more abstract threshold, or microsimulation like models yields a computationally tractable approach. Despite the relative simplicity, it nonetheless captures some of the nuance and sophistication of human decisions.

%Finishing up

The remainder of this paper proceeds to provide a brief review of the methodological context (\ref{sub:games_signals}), before outlining the model (\ref{sec:model}), and experiments (\ref{sec:method}), with selected results (\ref{sec:results}), then closing with a discussion contrasting the decision models (\ref{sec:conclusion}).

### 2014-07-16 Model.tex
Remove implicit insult to frequentist approach
Tweak notation to make Risk calculation clearer
Working on introduction

While in principle a wide variety of agent models are possible, given that decision rules operate on essentially the same information, and produce the same outputs, we limit ourselves here to four. The simplest is a lexicographic rule (1), motivated as in the spirit of a \ac{FFH} \citep{Gigerenzer2004} which uses only information about payoffs given actions; a Bayesian risk minimisation rule using the same information (2); a second Bayesian risk rule (3) which uses information about the underlying lottery; and a two-stage \ac{CPT} \cite{Hau2008} agent (4) which is identical with 3, but uses the \ac{CPT} decision rule from \cite{Tversky1992}. Hence, each successive decision model adds a layer of sophistication to the problem representation while retaining the same input-ouput characteristics.

@@

\begin{equation}
R_{w}(s_{w}) = \sum_{i\in s_{m}}\sum_{j\in \theta_{m}} -u_{w}(s_{w}, s_{m, i}, \theta_{w}, \theta_{m, j})p(j)p(i | s_{w})
\end{equation}

@@

The most complex decision rule used is \ac{CPT}, which attempts to reproduce a number of systematic deviations from rationality observed in humans. While \ac{CPT} has primarily been applied in the context of decisions from description, it has been modified to deal with decisions from experience by incorporating a first stage where probabilities are estimates from observations as in \cite{FoxCPT}. In this instance we replace the frequentist first stage with the equivalent Bayesian inference process.

Rather than the psychologically more interesting \ac{PT}, the \ac{CPT}
decision rule is used in this instance, because of the requirement
for women to evaluate more than two `prospects'.\footnote{A prospect in this instance is a paired outcome and probability, and the set of prospects
for an action hence define the outcome space.} \ac{CPT} introduces
the concept of a probability weighting function, which underweights
small probabilities, and overweights large ones in an effort to capture
the tendency of humans to treat high probability events as sure things,
and small probabilities as `never going to happen'. A number of different
weighting functions have been proposed, but in this instance the original
formulation by \citet{Tversky1992} is used. This distinguishes between
weighting for gains, and losses -

>>>>

While in principle a wide variety of agent models are possible, given that decision rules operate on essentially the same information, and produce the same outputs, we limit ourselves here to four. The simplest is a lexicographic rule (1), motivated as in the spirit of a \ac{FFH} \citep{Gigerenzer2004} which uses only information about payoffs given actions; a Bayesian risk minimisation rule using the same information (2); a second Bayesian risk rule (3) which uses information about the underlying lottery; and a two-stage \ac{CPT} \citep{Hau2008} agent (4) which is identical with 3, but uses the \ac{CPT} decision rule from \cite{Tversky1992}. Hence, each successive decision model adds a layer of sophistication to the problem representation while retaining the same input-ouput characteristics.

@@

\begin{equation}
R_{w}(s_{w}) = \sum_{i\in s_{m}}\sum_{j\in \theta_{m}} -u_{w}(s_{w}, s_{m, i}, \theta_{w}, \theta_{m, j})p(s_{m,i})p(i | s_{w})
\end{equation}

@@

The most complex decision rule used is \ac{CPT}, which attempts to reproduce a number of systematic deviations from rationality observed in humans. While \ac{CPT} has primarily been applied in the context of decisions from description, it has been modified to deal with decisions from experience by incorporating a first stage where probabilities are estimates from observations as in \cite{FoxCPT}. In this instance the Bayesian inference process fills the first stage role.

%Maybe cut this bit and make explanation richer?
Rather than the psychologically more interesting \ac{PT}, the \ac{CPT}
decision rule is used in this instance, because of the requirement
for women to evaluate more than two `prospects'.\footnote{A prospect in this instance is a paired outcome and probability, and the set of prospects
for an action hence define the outcome space.} \ac{CPT} introduces
the concept of a probability weighting function, which underweights
small probabilities, and overweights large ones in an effort to capture
the tendency of humans to treat high probability events as sure things,
and small probabilities as `never going to happen'. A number of different
weighting functions have been proposed, but in this instance the original
formulation by \citet{Tversky1992} is used. This distinguishes between
weighting for gains, and losses -
